What can GitHub and Googlecode Issues tell us about Exceptions in Android and Java?

### 2014-09-01 android-stacks.tex
improvements refs

%the mobile industry shipped over 315 million smartphones in the last quarter of 2013
%The pervasiveness of the World Wide Web and the astonishing increase 
%in the number of smartphones stimulated the creation of applications which extend the phones
% capabilities far beyond of the basic calls and textual messages. 
%f we look at global smartphone shipments the numbers are astonishng;
%the mobile industry shipped over 315 million smartphones in the last quarter of 2013 [ref-googleIO].
%Such increase on smartphones combined with the pervasiveness of the World Wide Web
%have stimulated the creation of applications which extend the phones capabilities 
%far beyond of the basic calls and textual messages.

@@

Our conjecture is that several Android applications may be suffering from similar 
fragilities on the exception-related code (of the applications themselves or 
the underlaying platform), and that such fragilities may be leading to
decrease the application robustness, leading for instance to the Uncaught Exception problem -
 which is one of the main causes of application crashes ~\cite{jo2004uncaught} ~\cite{zhang2012amplifying}.

@@

To do so we perform a post mortem analysis of the exception stack traces embedded on all issues
created on Android projects available in GitHub and Googlecode. Using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from issues across all Android projects on
GitHub and Googlecode. Overall, we analyzed all issues extracted from on 482 Android projects defined in
GitHub and 157 projects defined in Googlecode. Overall it corresponds to 74,256 issues (32,582 issues from GitHub projects 
and 41,674 issues from Googlecode projects) from which  6,005 stack traces were extracted and used as
 a source of information for our analysis. The stack trace analysis was augmented by additional information
extracted from JDK, Android SO and Android SDK and the applications using byte code and source code analysis.
 Manual inspection was used to leverage the understanding of exception stack traces and support further discussions and insights.

@@

   \item Undocumented checked exception signaled by native Android library. This finding points to a fragility 
on the way exceptions on JNI interface functions, which can lead to uncaught exceptions that are difficult do 
debug.

  \item  A multitude of programming mistakes were reported as the main causes of stack traces:
 XXX  of reported stack traces were caused by programming mistakes; and XXX of analyzed stack
 traces reported a java.lang.NullPointerException as root cause;

@@

\section{Background}

@@

Android is an open source platform for mobile devices based on Linux kernel,
which also contains  a core Java library, a set of libraries third-party libraries (e.g.  SQLite),
the Dalvik runtime environment, and the Application Framework (which provides 
higher-level services to the applications running on the platform) ~\cite{andguide}.
Android platform is mostly implemented in Java and uses JNI[ref] to access native C/C++ code. 
And as mentioned before, it reuses the embedded Java Exception model to detect and handle 
abnormal computation states.

@@

<AVALIAR MOVER ESTE STACK TRACES PARA UMA FIGURA QUE MOSTR COMO ELE É QUEBRADO...
PELO EXCEPTION MINER E AS INFORMAÇÕES VAO PARA O BANCO...>

@@

Nevertheless, several general guidelines have been proposed on how to use Java
exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
bloch2008effective} regardless the exception type used. Such guidelines do not focus on 
advocating any specific type but rather propose ways to effectively use each of them.
 We compiled a list of the guidelines related to exception types in Java, 
summarized below:\footnote{We also compiled guidelines related
to exception handling but they are out of the scope of this paper.}

@@

\section{Empirical Study Design}

@@

Then we performed further cleanup, inspecting the site of every Android
reporting at least one stack trace, to make sure that they represented real
mobile apps. During this clean up 106 apps were removed because they were either
example projects (i.e., toy projects) or tools to support Android development
(e.g. \textsf{selendroid}, \textsf{roboeletric} --- tools to support the testing of Android apps).
The filtered set consisted of 482 apps.

@@

\noindent\emph{Exception Classifier} To support a deeper investigation of the stack traces 
every exception defined on a stack trace need to be classified according to its type
(e.g. Error, Exception or RuntimeException). The exception classifier is based on 
bytecode analysis (based on Design Wizard~\cite{Brunet09}) that walks up the 
type hierarchy of a Java exception until it reaches a base exception type.
This module analyzed the exception hierarchy of all exceptions thrown on:
 the last version of SDK version 7, and Android Platform and OS (level 19).

\noindent\emph{Exception Signaler Classifier.} 
This module is responsible for classifying each signaler according 
to its origin (i.e. Android Plaform, Android Libcore, Aplication, Library). 
We downloaded the source code of Android Platform and OS (level 19)
and this module extracts the packages hierarchies (recursive list of 
of Java packages), that composes it. To discover the packages that compose 
each application, this module extracted the manifest files of each Android apps
 - which defines the main packages that compose the applications.
Then based on pattern matching between the signaler name and the packages 
 this module identifies the origin of the exception signalers.
The exceptions were considered from libraries if their packages were not defined 
on Android Platform, core libraries nor on applications.
Table ~ref{signalers} summarizes the signaler classification.

@@

\subsection{Manual Inspections}
This study was also based in a set of manual inspection tasks. Besides the 
 identification of Android projects details previously. 
Manual inspections were used discover the type of exceptions thrown by 
the applications and libraries we manually inspect the sourcecode or javadoc
 documentation of each of them. When the exception was removed from the 
last app version, the previous versions were analyzed. When the exception was
 not found on the previous versions, the exception was discarded.  Only 31 exceptions 
remained undefined - such exceptions were reported on 61 exception stack traces as 
presented in Table ~\ref{tab:typeroottab}.

@@

\section{The Study Results}

@@

After distilling the information available on the exception stack traces we could find 
the the exceptions commonly reported as the root causes of stack traces.
Table \ref{tab:toptenandroid} illustrates the top 10 root causes found in the study,
 ranked by the number of distinct project repositories on which they were reported. 
The table shows: the number of times each exception was reported as 
the root cause of a stack trace, the number of distinct projects reporting such causes on stack traces, 
and the percentages of such in relation to all reported exceptions and analyzed projects respectively.
It also presents the whether the method responsible for signaling the exceptions were
part of Android, the application itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers}

@@

%For each exception we could discover the number of times it was reported as the root caused on exception stack traces,
%the number of distinct projects on which such stacks were defined (and consequently its
%popularity) and the of times it was signaled by different signaler types. We
%could calculate the popularity of each root cause, considering the number of
%distinct projects they were found and the whole set of projects analyzed with at
%least one stack trace (482). Table~\ref{tab:toptenandroid} 
%and Figure~\ref{fig:androidsignaler} presents the mined data.

@@

    \hline
  \end{tabular}
\caption{Root Exceptions occurrences and popularity in repositories hosted in Googlecode (GC) and GitHub(GH).}
\label{tab:toptenandroid}
\end{table*}

@@

\emph{\textbf{The high number of NullPointerExceptions.}}
From this set the java.lang.NullPointerException was the most reported root cause; 1225 stack traces 
had a java.lang.NullPointerException  as root cause, which corresponds of 30,08\%  of 
all analyzed stack traces. Considering the number of distinct project repositories, 
the java.lang.NullPointerException, was reported on at least one issue of 99 projects 
(which represents 20,54\% of all projects reporting at least one stack trace).
 The NullPointerExceptions are mainly signaled inside Android platform and inside 
application code, although we also find
NullPointerExceptions being signaled from third-party libraries. Regarding
reusable code, there is no consensus whether it is a good or bad practice to 
re-throw a NullPointerException. Some prefer to encapsulate such an exception on
IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
NullPointerException makes the cause of the problem explicit and hence 
should not be wrapped.

The high prevalence of NullPointerExceptions is aligned with the findings other 
works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
al.~\cite{kim2013predicting} showed that in Eclipse bug report system 38\% the bugs 
related to exception handling were caused by NullPointerException; other works on robustness 
testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically 
detected bugs were due to NullPointerExceptions and implicitly-signaled of Java
environment (as the ones found in this study).In Section XX we provide further 
discussion on the NullPointerException problem, pointing 
to ways of how to mitigate it.

\emph{\textbf{Direct instances of RuntimeExceptions being thrown.}}
From Table ~\ref{tab:toptenandroid} we can observe that direct instances
java.lang.RuntimeException were thrown in 18,88\% of Android repositories, and
most of such exceptions  were  thrown by the Android platform (167 out of 232).

@@

Such categorization was based known sources of faults mobile development: concurrency [INCLUDEREF]
, backward compatibility  [INCLUDEREF], security   [INCLUDEREF] and resource management (IO, Memory, Batery)  [INCLUDEREF].
We added to this set categories to represent the exceptions caused by programming mistakes,
 and general Java exceptions (whose types not convey any specific cause of failure), and 
exceptions that represent specific  exceptional situations of apps, frameworks or libraries.
The list of categories is illustrated in Table~\ref{tab:categories}.

@@

The code snippet in Figure X illustrate this scenario.
 The java/lang/Class declares a native method named getDeclaredMethods. 
The Java-side declaration of of this method does not have any throws clause, 
leading programmers and the compiler to think that no checked exceptions can be thrown.

@@

Figure X presents a code snippet of ActivityThreat which is responsible for (among other things)
calling the Activities onCreate method. Hence, as we can see from lines 4-8,
 in case any exception is thrown by an application Activity during its creation, 
it is wrapped in a general Runtime exception and re-thrown.

@@

We could observe that several types of exceptions were encapsulated in 
instances of Runtime. Table X shows some kinds of errors encapsulated a
RuntimeException.

%\begin{table}
%\centering
%\begin{tabular}{ll}
% \bfseries{Wrapper type - Root cause type} & \bfseries{\#} \\
%    \hline
%java.lang.RuntimeException - java.lang.OutOfMemoryError & 30\\
%java.lang.RuntimeException - java.lang.NoClassDefFoundError & 6\\
%java.lang.RuntimeException - java.lang.UnsatisfiedLinkError) & 5\\
%java.lang.RuntimeException -  java.lang.StackOverflowError)  & 4\\
%\hline
%  \end{tabular}
%\caption{Elements responsible for the wrapping.}
%\label{tab:wrapping01}
%\end{table}

@@

 %\begin{tabular}{|p{7.9cm} p{0.1cm}|}
\begin{tabular}{ll}
    \hline
 id & \bfseries{Runtime-Error Wrapping Examples}    \\  %\bfseries{\#}
    \hline
1 & java.lang.RuntimeException - java.lang.OutOfMemoryError  \\ %30
2 &java.lang.RuntimeException - java.lang.NoClassDefFoundError \\ %6
3 &java.lang.RuntimeException - java.lang.UnsatisfiedLinkError)  \\ %5
4& java.lang.RuntimeException -  java.lang.StackOverflowError)   \\ %4
\hline
 & \bfseries{Checked-Runtime Wrapping Examples}   \\
 \hline
5& ...android.robospice.CacheSavingException - java.lang.ClassCastException   \\ %6
6&java.lang.reflect.InvocationTargetException - java.lang.NullPointerException    \\ %3
7&java.lang.reflect.InvocationTargetException - ...ResourcesdNotFoundException  \\ %2
8&java.sql.SQLException - android.database.sqlite.SQLiteDiskIOException   \\ %2
 \hline
& \bfseries{Checked-Error Wrapping Examples}  \\
 \hline
9&java.lang.reflect.InvocationTargetException - java.lang.OutOfMemoryError  \\ %5
10&java.lang.Exception - java.lang.OutOfMemoryError   \\ %1
11&java.lang.reflect.InvocationTargetException - java.lang.NoSuchMethodError 	 \\ %1
12&java.lang.reflect.InvocationTargetException - java.lang.UnsatisfiedLinkError	 \\ %1
\hline
& \bfseries{Error-Checked Wrapping Examples}  \\
 \hline
13&java.lang.NoClassDefFoundError - java.lang.ClassNotFoundException   \\ %4
14&java.lang.AssertionError - javax.crypto.ShortBufferException)   \\ %3
\hline
& \bfseries{Error-Runtime Wrapping Examples}    \\
 \hline
15&java.lang.ExceptionInInitializerError - java.lang.NullPointerException   \\ %9
16&java.lang.ExceptionInInitializerError - java.lang.IllegalArgumentException 	 \\ %2
 \hline
  \end{tabular}
\caption{Examples of Cross-type wrappings}
\label{tab:wrapping01}
\end{table}

@@

\section{Futher Discussions}

\begin{figure*} \centering \includegraphics[scale=0.55]{codeexamples.pdf}
\caption{Code snippets illustrating fragilities on the exception-related code} \label{fig:exchier} \end{figure*}

@@

This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations;
(ii) inter-method null pointer analysis tools such as the one proposed by [DANIELE-RELATED-WORK];
or (iii) whether language designs which avoid null pointers, such 
as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
or computations that may fail) could improve the robustness of Java programs.

@@

\subsection{Actionable Advices - The Unwritten rules of Android development}

@@

\section{Threats to Validity}

@@

\section{Related Work}

@@

\section{Conclusion}

In this paper we present an exploratory study in which we mined the stack 
traces embedded in all issues of Java projects available GitHub.
Furthermore, we analyzed in more detail the
stack traces reported on a subset of 482 Android projects. In this study, the information extracted 
from stack traces was used in combination with source code and bytecode analysis to 
pin-point patterns of exception use and misuse in Java.
Certain patterns of exception
(mis)-use were consistently detected such as unexpected wrappings (e.g., Errors
being wrapped in checked exceptions), revealing that  the Java hybrid exception
model is not fully used according to its purpose; undocumented runtime
exceptions signaled by third party code (which makes it almost impossible for
library clients to protect against such exceptions); 
and a high prevalence of
\texttt{java.lang} exceptions reported on issues (representing approximately 50\% of the
analyzed issues). Such (mis)uses detected in our study can negatively affect the 
robustness of current Java software systems. Our results call for  
 developing tools support to help developers dealing with such problems and 
improving the design of the exception handling mechanism in Java.

>>>>

%the mobile industry shipped over 315 million smartphones in the last quarter of 2013
%The pervasiveness of the World Wide Web and the astonishing increase 
%in the number of smartphones stimulated the creation of applications which extend the phones
% capabilities far beyond of the basic calls and textual messages. 
%f we look at global smartphone shipments the numbers are astonishng;
%the mobile industry shipped over 315 million smartphones in the last quarter of 2013 [ref-googleIO].
%Such increase on smartphones combined with the pervasiveness of the World Wide Web
%have stimulated the creation of applications which extend the phones caVALIAR MOVER ESTE STACK TRACES PARA UMA FIGURA QUE MOSTR COMO ELE QUEBRADO... PELO EXCEPTION Mpabilities 
%far beyond of the basic calls and textual messages.

@@

Our conjecture is that several Android applications may be suffering from similar 
fragilities on the exception-related code (of the applications themselves or 
the underlaying platform), and that such fragilities may be leading to
decrease the application robustness, leading for instance to the Uncaught Exception problem -
 which is one of the main causes of application crashes ~\cite{jo2004uncaught, zhang2012amplifying}.

@@

To do so we perform a post mortem analysis of the exception stack traces embedded on all issues
created on Android projects available in GitHub and Googlecode. Using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from issues across all Android projects on
GitHub and Googlecode. Overall, we analyzed all issues extracted from on 482 Android projects defined in
GitHub and 157 projects defined in Googlecode. Overall it corresponds to 74,256 issues (32,582 issues from GitHub projects 
and 41,674 issues from Googlecode projects) from which  6,005 stack traces were extracted and used as
 a source of information for our analysis. The stack trace analysis was augmented by additional information
extracted from JDK, Android Plaform and the applications using byte code and source code analysis.
 Manual inspection was used to leverage the understanding of exception stack traces and support further discussions and insights.

@@

   \item Undocumented checked exception signaled by native Android library. This finding points to a fragility 
on the way exceptions flow from JNI interface functions, which can lead to uncaught exceptions that are difficult do 
debug.

  \item  A multitude of programming mistakes were reported as the main causes of stack traces:
 27,71\% of the analyzed exception stack traces reported a java.lang.NullPointerException the root cause;
and  51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
was the root cause.

@@

\section{Background}
\label{sec:back}

@@

Android is an open source platform for mobile devices based on Linux kernel,
which also contains  a core Java library, a set of libraries third-party libraries (e.g.  SQLite),
the Dalvik runtime environment, and the Application Framework (which provides 
higher-level services to the applications running on the platform) ~\cite{andguide}.
Android platform is mostly implemented in Java and uses Java Native Invocation (JNI) to access native C/C++ code. 
And as mentioned before, it reuses the embedded Java Exception model to detect and handle 
abnormal computation states.

@@

%<AVALIAR MOVER ESTE STACK TRACES PARA UMA FIGURA QUE MOSTR COMO ELE É QUEBRADO...
%PELO EXCEPTION MINER E AS INFORMAÇÕES VAO PARA O BANCO...>

@@

Nevertheless, several general guidelines have been proposed on how to use Java
exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
bloch2008effective}. Such guidelines do not focus on 
advocating any specific exception type, but rather propose ways to effectively use each of them.
 We compiled a list of the guidelines related to exception types in Java, 
summarized below:\footnote{We also compiled guidelines related
to exception handling but they are out of the scope of this paper.}

@@

\section{Empirical Study Design}
\label{sec:study}

@@

Then we performed further cleanup, inspecting the site of every Android
reporting at least one stack trace, to make sure that they represented real
mobile apps. During this clean up 106 apps were removed because they were either
example projects (i.e., toy projects) or tools to support Android development
(e.g. Selendroid, Roboeletric - tools to support the testing of Android apps).
The filtered set consisted of 482 apps.

@@

\noindent\emph{Exception Classifier} To support a deeper investigation of the stack traces 
every exception defined on a stack trace need to be classified according to its type
(e.g. Error, Exception or RuntimeException). The exception classifier is based on 
bytecode analysis (based on Design Wizard~\cite{Brunet09}) that walks up the 
type hierarchy of a Java exception until it reaches a base exception type.
This module analyzed the exception hierarchy of all exceptions thrown on:
  Android Platform (API level 19), and JDK (version 6).

\noindent\emph{Exception Signaler Classifier.} 
This module is responsible for classifying each signaler according 
to its origin (i.e. Android Plaform, Android Libcore, Aplication, Library). 
We downloaded the source code of Android Platform (API level 19)
and this module extracts the recursive list of 
of Java packages that composes it. To discover the packages that compose 
the previous released of the API we performed a manual inspection on the Javadoc
documentation available online for each API level ~\cite{apidocs}.
To discover the packages composing the each application, this module 
extracted the manifest files of each Android app
 - which defines the main packages that compose the applications.
Then based on pattern matching between the signaler name and the packages 
 this module identifies the origin of the exception signalers.
The exceptions were considered from libraries if their packages were not defined 
on Android Platform, core libraries nor on applications.
Table ~\ref{signalers} summarizes the signaler classification.

@@

\subsection{Manual Inspections}
As mentioned before, this study was also based in a set of manual inspection tasks. Besides the 
 (i) identifying of Android projects, and (ii) the finding packages that compose the 
previous Android APIs leveles, manual inspections were also used discover the type of associated to some 
reported on stack traces. Such exceptions could not be automaticaly identified 
because they were not defined on the last version of the application,
or were defined on libraries not fount among projects resources.
In such cases we manually inspect the sourcecode or javadoc
 documentation available online of each of them. 
When the exception was not found on the previous versions, the exception where classified 
in our study as "Undefined".  Only 31 exceptions 
remained undefined - such exceptions were reported on 61 exception stack traces as 
presented in Table ~\ref{tab:typeroottab}.

@@

\section{The Study Results}
\label{sec:result}

@@

After distilling the information available on the exception stack traces we could find 
the the exceptions commonly reported as the root causes of stack traces.
Table \ref{tab:topten} illustrates the top 10 root causes found in the study,
 ranked by the number of distinct project repositories on which they were reported. 
The table shows: the number of times each exception was reported as 
the root cause of a stack trace, the number of distinct projects reporting such causes on stack traces, 
and the percentages of such in relation to all reported exceptions and analyzed projects respectively.
It also presents the whether the method responsible for signaling the exceptions were
part of Android, the application itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers}

@@

%For each exception we could discover the number of times it was reported as the root caused on exception stack traces,
%the number of distinct projects on which such stacks were defined (and consequently its
%popularity) and the of times it was signaled by different signaler types. We
%could calculate the popularity of each root cause, considering the number of
%distinct projects they were found and the whole set of projects analyzed with at
%least one stack trace (482). Table~\ref{tab:topten} 
%and Figure~\ref{fig:androidsignaler} presents the mined data.

@@

    \hline
  \end{tabular}
\caption{Root Exceptions occurrences and popularity in repositories hosted in Googlecode (GC) and GitHub(GH).}
\label{tab:topten}
\end{table*}

@@

\emph{\textbf{The high number of NullPointerExceptions.}}
From this set the java.lang.NullPointerException was the most reported root cause; 27,71\% 
of the analyzed exception stack traces reported a java.lang.NullPointerException the root cause. 
Considering the frequencly of NullPointerException across projects, we could observe that
 51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
was the root cause.
 The NullPointerExceptions are mainly signaled inside the applicaiton code and the Android platform,
 although we also find NullPointerExceptions being signaled from third-party libraries. Regarding
reusable code, there is no consensus whether it is a good or bad practice to 
re-throw a NullPointerException. Some prefer to encapsulate such an exception on
IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
NullPointerException makes the cause of the problem explicit and hence 
should not be wrapped.

The high prevalence of NullPointerExceptions is aligned with the findings of other 
works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
al.~\cite{kim2013predicting} showed that in Eclipse bug report system 38\% the bugs 
related to exception handling were caused by NullPointerException; other works on robustness 
testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically 
detected bugs were due to NullPointerExceptions 
and exceptions implicitly-signaled by Java environment due to programming mistakes or resource limitations
 (as the ones found in this study).In Section ~\ref{sec:disc}  we provide further 
discussion on the NullPointerException problem, pointing 
to ways of how to mitigate it.

\emph{\textbf{Direct instances of RuntimeExceptions being thrown.}}
From Table ~\ref{tab:topten} we can observe that direct instances
java.lang.RuntimeException were thrown in 18,88\% of Android repositories, and
most of such exceptions  were  thrown by the Android platform (167 out of 232).

@@

Such categorization was based known sources of faults mobile development: concurrency ~\cite{ama2012}
, backward compatibility  ~\cite{McDon13}, security   ~\cite{enck2011study} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
We added to this set of categories other three categories: one to represent the exceptions caused by programming mistakes;
  another to represent general Java exceptions (e.g., java.lang.Exception) - whose types do not convey any specific cause of failure; and a category to
represent specific  exceptional situations of apps, frameworks and libraries.
The list of categories is illustrated in Table~\ref{tab:categories}.

@@

The code snippet in Figure ~\ref{fig:snippets} (d) illustrate this scenario.
 The java/lang/Class declares a native method named getDeclaredMethods. 
The Java-side declaration of of this method does not have any throws clause, 
leading programmers and the compiler to think that no checked exceptions can be thrown.

@@

Figure ~\ref{fig:snippets} (a) presents a code snippet of ActivityThreat which is responsible for (among other things)
calling the Activities onCreate method. Hence, as we can see from lines 4-8,
 in case any exception is thrown by an application Activity during its creation, 
it is wrapped in a general Runtime exception and re-thrown.

@@

We could observe that several types of exceptions were encapsulated in 
instances of Runtime such as java.lang.OutOfMemoryError (30), java.lang.NoClassDefFoundError (6), 
java.lang.UnsatisfiedLinkError (5), java.lang.StackOverflowErro (4), as illustrated in Table ~\ref{tab:exampeswrap}.

@@

 %\begin{tabular}{|p{7.9cm} p{0.1cm}|}
\begin{tabular}{ll}
    \hline
 id & \bfseries{Runtime-Error Wrapping Examples}    \\  %\bfseries{\#}
    \hline
1 & java.lang.RuntimeException - java.lang.OutOfMemoryError  \\ %30
2 &java.lang.RuntimeException - java.lang.NoClassDefFoundError \\ %6
3 &java.lang.RuntimeException - java.lang.UnsatisfiedLinkError)  \\ %5
4& java.lang.RuntimeException -  java.lang.StackOverflowError)   \\ %4
\hline
 & \bfseries{Checked-Runtime Wrapping Examples}   \\
 \hline
5& ...android.robospice.CacheSavingException - java.lang.ClassCastException   \\ %6
6&java.lang.reflect.InvocationTargetException - java.lang.NullPointerException    \\ %3
7&java.lang.reflect.InvocationTargetException - ...ResourcesdNotFoundException  \\ %2
8&java.sql.SQLException - android.database.sqlite.SQLiteDiskIOException   \\ %2
 \hline
& \bfseries{Checked-Error Wrapping Examples}  \\
 \hline
9&java.lang.reflect.InvocationTargetException - java.lang.OutOfMemoryError  \\ %5
10&java.lang.Exception - java.lang.OutOfMemoryError   \\ %1
11&java.lang.reflect.InvocationTargetException - java.lang.NoSuchMethodError 	 \\ %1
12&java.lang.reflect.InvocationTargetException - java.lang.UnsatisfiedLinkError	 \\ %1
\hline
& \bfseries{Error-Checked Wrapping Examples}  \\
 \hline
13&java.lang.NoClassDefFoundError - java.lang.ClassNotFoundException   \\ %4
14&java.lang.AssertionError - javax.crypto.ShortBufferException)   \\ %3
\hline
& \bfseries{Error-Runtime Wrapping Examples}    \\
 \hline
15&java.lang.ExceptionInInitializerError - java.lang.NullPointerException   \\ %9
16&java.lang.ExceptionInInitializerError - java.lang.IllegalArgumentException 	 \\ %2
 \hline
  \end{tabular}
\caption{Examples of Cross-type wrappings}
\label{tab:exampeswrap}
\end{table}

@@

\section{Further Discussions}
\label{sec:disc}

\begin{figure*} \centering \includegraphics[scale=0.55]{codeexamples.pdf}
\caption{Code snippets illustrating fragilities on the exception-related code} \label{fig:snippets} \end{figure*}

@@

This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations;
(ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
or (iii) whether language designs which avoid null pointers, such 
as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
or computations that may fail) could improve the robustness of Java programs.

@@

\subsection{Actionable Advices - The Unwritten rules of Android development}
to do if we have space...

@@

\section{Threats to Validity}
\label{sec:threats}

@@

\section{Related Work}
\label{sec:rele}

@@

\section{Conclusion}
\label{sec:conc}

In this paper we present an exploratory study in which we mined the stack 
traces embedded in all issues defined in 482 Android projects hosted in Github and 
157 projects hosted in Googlecode.

In this study, the information extracted from stack traces was used in combination 
with source code and bytecode analysis to pin-point fragilities on the exception-related
code. 
This study  consistently identifies common characteristics of reported 
exception stacks traces - e.g., the NullPointerException was the reported cause 
on 30,08\% of all analyzed exception stack traces, and was mentioned in approximately 
50\% of all projects analyzed. Moreover,  the cross-project analysis of exception stack traces 
enabled to consistently reveal  fragilities of the exception-related code: 
(i) undocumented runtime exceptions signaled by third party code,
(ii) undocumented checked exceptions thrown by Android native code
(iii) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
revealing that Java hybrid exception model is not fully used according to its purpose.
Such fragilities detected in our study can negatively affect the 
robustness of not only Android application but also Java-based 
which can deal with the same fragilities pointed in thid work. 
Our results calls the attention for the need of tool and language support to help 
developers when dealing with such problems.
